---
title: "Evaluating Recruiting Channels"
author: "Ben Teusch"
date: "September 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
##### <span style="color:blue"> I believe my audience is people who have downloaded R, and are comfortable with at least copying code and running it. No actual knowledge of R is expected, but a willingness to learn on your own is, because I'm not covering everything here. Analysts trying to move into R should find this helpful, and some R users trying to break into people analytics will find the R simple, but hopefully will find the applications interesting.</span>

### Load in data
I'm going to load in the data using `read.csv`. This function can be used to read in a .csv file from your computer, or from the Internet. Here, I'll load in the data I've saved on my github site, and store it as a `data.frame` called `mydata`.
```{r}
mydata <- read.csv("https://raw.githubusercontent.com/teuschb/hr_data/master/datasets/recruitment_evaluation_data.csv")
```

### Examine/explore data
Whenever I have a new dataset, I like to see what I'm working with. I tend to use `str()` and `summary()`. The  `str()` function shows the structure of your data: the column names and how many there are, and what type of data each one is. The `summary()` function will show means and quadriles of each column, as well as how many observations are missing. If you're used to working with spreadsheets, `View(mydata)` will show you a spreadsheet with your data (note the capital **V**).

```{r}
str(mydata)
```

I see there is a column called HireSource that shows which recruiting channel our hires came from. Instead of using `summary()` on the entire dataset, let's use it on just the HireSource column. The `$` operator is one way to specify a column in R.

```{r}
summary(mydata$HireSource)
```

This shows the number of hires that came from the four broad hiring channels we have, as well as `NA`, which R uses to denote missing data. It can be helpful to understand why data is missing. In this case, I know that our ATS only started tracking hire data 6 years ago, so anybody hired over 6 years ago won't have this data. I can check that by calling `summary()` on `mydata$TotalWorkingYears`, subset by only the people whose `HireSource` information is missing: `is.na(mydata$HireSource)`.

```{r}
summary(mydata$TotalWorkingYears[is.na(mydata$HireSource)])
```

##### <span style="color:blue">Would also be good here to see if the recruiting channel has the same distribution across departments</span>


### Analyze data
Since I'm trying to determine which of these hiring channels is best, I need some way to compare them. Hopefully I know what is most important in my company for recruiting success, but if I want to quickly see all the data I could use, another trick is the `names()` function, which lists all the columns in my data very cleanly.

```{r}
names(mydata)
```

Scanning over this I see `Attrition`, `PerformanceRating`, and `SalesRating` as possible outcome variables (or measures of success). 

#### Attrition
Let's start with attrition. If employees from one of the recruiting channels turn over less, I might want to do more recruiting there in the future. Let's look at the data:

```{r}
summary(mydata$Attrition)
```

This is very simple attrition data. I only have one time period, and I don't know whether the attrition was voluntary or involuntary, or whether the employee retired, or lots of other things I might wonder about. For this example, however, I'll use what I have and compare the attrition across the recruiting channels.

In Excel, I might use a pivot table for this. There are several ways to do it in R. My favorite requires installing and loading a `package`, which is a separate set of functions contributed by the R community, but not part of the base R installation.

In this case, the package is `psych`, so before running the next code block, you'll need to type `install.packages("psych")` into your R console to download and install the `psych` package to your computer. The `describeBy()` function in this next block is part of the `psych` package. Before we can use it, I have to load the package into my current session using `library()`.

```{r}
library(psych)
# describe attrition by hiresource, and put the output into a matrix
describeBy(mydata$Attrition, mydata$HireSource, mat = TRUE)  
```

I like `describeBy()` because the information is output close together, so it is easy to compare. I don't always care about range or skew, but if I did, it's all there. Here I'm most curious in the mean. Is the average attrition the same for each recruiting channel, or are there differences?

<span style="color:blue">what graphs should I be using?</span>

Visually comparing the means, it looks like employees hired by a search firm have higher attrition, and referral hires have lower attrition. Is that difference significant? We can check by using a t-test, which will help us know if the difference in attrition between the two groups is real, or just due to variation in the data.

##### <span style="color:blue">data will be fixed so search firm as higher attrition and referral has lower</span>

Using the function `t.test()`, I can compare attrition from employees whose `HireSource` is Search Firm with employees whose `HireSource` is Referral. See if you understand how I use the `$` and `[` operators to tell R to look at those cuts of data.

```{r}
t.test(mydata$Attrition[mydata$HireSource == "Search Firm"], mydata$Attrition[mydata$HireSource == "Referral"])
```

<span style="color:blue">regression</span>

```{r}
att_mod <- glm(Attrition ~ HireSource + Department + BusinessTravel + OverTime + YearsAtCompany +
                TrainingTimesLastYear + WorkLifeBalance,
               family = binomial(link = 'logit'), data = mydata)
summary(att_mod)
exp(summary(att_mod)$coefficients)
```



#### Performance Ratings

<span style="color:blue">same as attrition, basically</span>